\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

\subsubsection{Contributions of this paper}
Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be performed in a gridless and scalable manner. In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in way that is amenable for the aforesaid generalized proximal recursions.

@@

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular,
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous. Fig. \ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.

@@

\begin{algorithm}[t]
\caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
\label{algo:KineticProx}
\begin{algorithmic}[1]
\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, L$}
\For{$i=1$ to $N$}
\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
\For{$j=1$ to $N$}
\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
\EndFor
\EndFor
\State $\bm{\Gamma}_{k} \gets \exp\left(-\bm{C}_{k}/2\varepsilon\right)$ \Comment{elementwise exponential}
\State $\bm{z}_{0} \gets {\rm{rand}}_{N\times 1}$ \Comment{random vector of size $N\times 1$}
\State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\bm{y} \gets \left[\tilde{\bm{\varrho}}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\ell = 1$
\While{$\ell\leq L$}
\State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
\State $\bm{y}(:,\ell+1) \gets \tilde{\bm{\varrho}}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
\If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
\State break
\Else
\State $\ell \gets \ell + 1$
\EndIf
\EndWhile\\
\Return $\tilde{\bm{\varrho}}_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
\EndProcedure	
\end{algorithmic}
\end{algorithm}

@@

The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$. Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples. The parameters $\delta$ and $L$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.

@@

With step size $h=10^{-3}$, we discretized time $t\in[0,1]$. Using the algorithm detailed in Sec. \ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$. In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.

@@

\begin{table}[t]
\centering
\begin{tabular}{| l | l |} 
\hline
Parameter description & Values\\
\hline\hline
&\\
nominal frequency & $f_{0} = 60$ Hz\\
&\\
inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
& \\
damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
& \\
diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
& \\
tangent of phase shift & $\tan\varphi_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0,0.25]\right)& \text{for}\; i\neq j
\end{cases}$\\
& \\
effective power input & $P_{i}\in{\rm{Unif}}\left([0,10]\right)$\\
& \\
coupling coefficient & $k_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j	
\end{cases}$\\
& \\
\hline
\end{tabular}
\caption{Parameters used for the numerical simulation in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
\label{table:ParamUncertaintiesSynthetic}
\vspace*{-0.35in}
\end{table}

\subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
To highlight the scalability of the proposed method, we next consider a power network with $n=50$ generators, and propagate the transient joint state PDFs supported over the 100 dimensional state space $\mathbb{T}^{50}\times\mathbb{R}^{50}$. We take the initial joint state PDF at $t=0$ as
% \begin{align}
% \rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
% 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
% \label{rho0synthetic}	
% \end{align}
\begin{multline}
\rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right) \\
	\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
\label{rho0synthetic}	
\end{multline}
and following \cite[Sec. 5]{dorfler2012synchronization}, randomly generate the parameters as in Table \ref{table:ParamUncertaintiesSynthetic}. These parameter ranges are consistent with the same found in \cite{sauerpai1998,andersonBook1977,kundurBook1994}.

@@

With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the procedure detailed in Sec. \ref{sec:ProxAlgo} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. As in Sec. \ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$ in Algorithm \ref{algo:KineticProx}.

>>>>

\subsubsection{Contributions of this paper}
Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be performed in a gridless and scalable manner. In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in a way that is amenable for the aforesaid generalized proximal recursions.

@@

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular\footnote{$L^{1}$ denotes the Lebesgue space of absolutely integrable functions.},
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous. Fig. \ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.

@@

\begin{algorithm}[t]
\caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
\label{algo:KineticProx}
\begin{algorithmic}[1]
\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
\For{$i=1$ to $N$}
\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
\For{$j=1$ to $N$}
\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
\EndFor
\EndFor
\State $\bm{\Gamma}_{k} \gets \exp\left(-\bm{C}_{k}/2\varepsilon\right)$ \Comment{elementwise exponential}
\State $\bm{z}_{0} \gets {\rm{rand}}_{N\times 1}$ \Comment{random vector of size $N\times 1$}
\State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\bm{y} \gets \left[\tilde{\bm{\varrho}}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\ell = 1$
\While{$\ell\leq \ell_{\max}$}
\State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
\State $\bm{y}(:,\ell+1) \gets \tilde{\bm{\varrho}}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
\If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
\State break
\Else
\State $\ell \gets \ell + 1$
\EndIf
\EndWhile\\
\Return $\tilde{\bm{\varrho}}_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
\EndProcedure	
\end{algorithmic}
\end{algorithm}

@@

The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$. Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples. The parameters $\delta$ and $\ell_{\max}$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.

@@

With step size $h=10^{-3}$, we discretized time $t\in[0,1]$. Using the algorithm detailed in Sec. \ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$. In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $\ell_{\max}=300$.

@@

\begin{table}[t]
\centering
\begin{tabular}{| l | l |} 
\hline
Parameter description & Values\\
\hline\hline
&\\
nominal frequency & $f_{0} = 60$ Hz\\
&\\
inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
& \\
damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
& \\
diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
& \\
tangent of phase shift & $\tan\varphi_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0,0.25]\right)& \text{for}\; i\neq j
\end{cases}$\\
& \\
effective power input & $P_{i}\in{\rm{Unif}}\left([0,10]\right)$\\
& \\
coupling coefficient & $k_{ij}\begin{cases}
=0 & \text{for}\; i=j\\
\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j	
\end{cases}$\\
& \\
\hline
\end{tabular}
\vspace*{0.1in}
\caption{Parameters used for the numerical simulation in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
\label{table:ParamUncertaintiesSynthetic}
\vspace*{-0.35in}
\end{table}

\subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
To highlight the scalability of the proposed method, we next consider a power network with $n=50$ generators, and propagate the transient joint state PDFs supported over the 100 dimensional state space $\mathbb{T}^{50}\times\mathbb{R}^{50}$. We take the initial joint state PDF at $t=0$ as
% \begin{align}
% \rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
% 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
% \label{rho0synthetic}	
% \end{align}
\begin{multline}
\rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right) \\
	\qquad\times {\rm{Unif}}\left(\left([-0.1,0.1]\;\text{rad/s}\right)^{n}\right),
\label{rho0synthetic}	
\end{multline}
and following \cite[Sec. 5]{dorfler2012synchronization}, randomly generate the parameters as in Table \ref{table:ParamUncertaintiesSynthetic}. These parameter ranges are consistent with the same found in \cite{sauerpai1998,andersonBook1977,kundurBook1994}.

@@

With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the procedure detailed in Sec. \ref{sec:ProxAlgo} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. As in Sec. \ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $\ell_{\max} = 300$ in Algorithm \ref{algo:KineticProx}.
